---
title: "From 16S Amplicon Raw Data to ASVs (Amplicon Sequence Variants)"
author: "IMC Bioinformatics"
date: "`r format(Sys.time(), '%d %B, %Y')`"
params:
   rmd: "qc_report.Rmd"
output:
html_document:
    theme: flatly
    highlight: tango
    toc: yes
    number_sections: yes
    toc_depth: 4
    toc_float:
      collapsed: no
    smooth_scroll: yes
---

<br>
<br>

```{r needed packages,echo=F,include=F}

suppressMessages(library(DT))
suppressMessages(library(dplyr))
suppressMessages(library(ggplot2))
suppressMessages(library(psadd))
suppressMessages(library(phyloseq))
suppressMessages(library(dada2))
suppressMessages(library(RColorBrewer))
suppressMessages(library(limma))
suppressMessages(library(tidyverse))
suppressMessages(library(reshape2))
suppressMessages(library(kableExtra))
suppressMessages(library(waterfalls))
suppressMessages(library(ggpubr))


Sys.setenv(RSTUDIO_PANDOC=snakemake@config[["pandoc"]])


```

# **Investigating reads**

<br>

## Overall read quality plots {.tabset .tabset-fade}

<br>

### Read quality raw 

<br>

```{r,results='asis', echo=F}

files <- list.files(path = snakemake@params[["quality"]], pattern = "raw", full.names = TRUE)

for (f in files) {
  cat(paste0("![](", f, "){width=50%}"))
 }

```

<br>
<br>
<br>
<br>
<br>


### Read Quality after read trimming

<br>

```{r, results='asis',echo=F}

files <- list.files(path = snakemake@params[["quality"]], pattern = "afterQC", full.names =TRUE)

for (f in files) {
  cat(paste0("![](", f, "){width=50%}"))
 }
```

<br>
<br>
<br>
<br>
<br>

### Read Quality after read filtering

<br>

```{r, results='asis', echo=F}

files <- list.files(path = snakemake@params[["quality"]], pattern = "afterdada2", full.names =TRUE)

for (f in files) {
  cat(paste0("![](", f, "){width=50%}"))
 }
```

<br>
<br>
<br>
<br>
<br>



```{r read length distribution plot execution, echo=F}
nsample<- snakemake@config[["random_n"]]

x = seq_along(1:nsample)

```

## Read length distribution for 5 random samples in the run {.tabset .tabset-fade}

<br>


<br>

```{r, results='asis', echo=F,eval =ifelse(1 %in% x, TRUE, FALSE)}

cat("### Sample 1\n")

files <- list.files(path = snakemake@params[["length_distribution"]], pattern = "S1", full.names =TRUE)

for (f in files) {
  cat(paste0("![](", f, ")\n"))
 }

```

<br>
<br>
<br>
<br>
<br>


```{r, results='asis', echo=F,eval =ifelse(2 %in% x, TRUE, FALSE)}

cat("### Sample 2\n")

files <- list.files(path = snakemake@params[["length_distribution"]], pattern = "S2", full.names =TRUE)

for (f in files) {
  cat(paste0("![](", f, ")\n"))
 }
```

<br>
<br>
<br>
<br>
<br>

```{r, results='asis', echo=F,eval =ifelse(3 %in% x, TRUE, FALSE)}

cat("### Sample 3\n")

files <- list.files(path = snakemake@params[["length_distribution"]], pattern = "S3", full.names =TRUE)

for (f in files) {
  cat(paste0("![](", f, ")\n"))
 }
```

<br>
<br>
<br>
<br>
<br>


```{r, results='asis', echo=F,eval =ifelse(4 %in% x, TRUE, FALSE)}

cat("### Sample 4\n")

files <- list.files(path = snakemake@params[["length_distribution"]], pattern = "S4", full.names =TRUE)

for (f in files) {
  cat(paste0("![](", f, ")\n"))
 }
```

<br>
<br>
<br>
<br>
<br>


```{r, results='asis', echo=F,eval =ifelse(5 %in% x, TRUE, FALSE)}

cat("### Sample 5\n")

files <- list.files(path = snakemake@params[["length_distribution"]], pattern = "S5", full.names =TRUE)

for (f in files) {
  cat(paste0("![](", f, ")\n"))
 }
```

<br>
<br>
<br>
<br>
<br>

```{r,echo=F,messages=FALSE, warning=FALSE}

reads<-read.csv(snakemake@params[["Nread"]],sep="\t") %>% filter(!grepl('Undetermined|undetermined', file))

```

## Reads count change throughout the dada2 pipeline in all samples

<br>

```{r table merge,echo=F,messages=FALSE, warning=FALSE,include=T,fig.width=9,fig.height=7}
reads1<-reads

colnames(reads1)[1:7]<-c("Sample", "Raw", "Quality_Trimmed", "Filtered", "Denoised", "Merged", "Cleaned")

reads1 <- reads1 %>% mutate_all(~ifelse(is.na(.), 0, .))

read_loss<-data.frame(step=names(colSums(reads1[,-c(1, 8)])),count=colSums(reads1[,-c(1, 8)]))
read_loss<-read_loss[-6,]

# Create a color palette for the bars (you can customize these colors)
colors <- c("Raw" = "#f5f0e1", "Quality_Trimmed" = "#ff6e40",
            "Filtered" = "#ff6e40", "Denoised" = "#ff6e40",
            "Merged" = "#ff6e40")



# Create a new data frame with the "step" and "loss" columns
new_data <- data.frame(
  step = read_loss$step,
  loss = c(read_loss$count[1],
           diff(read_loss$count)[1:4])) %>%
mutate(label = paste(step, "\n(n=", loss, ")"))

waterfall(new_data, calc_total = TRUE,
          fill_by_sign = F,
          rect_text_size = 1.5,
          fill_colours = colors,
          total_rect_color = "#ffc13b",
          total_rect_text_color = "black")+
  theme_bw()+xlab("Read Counts")+ylab("Processing steps")+ggtitle("Read count change in each step")+
  theme(axis.text.x = element_text(angle = 45, size = 14, colour = "black",  hjust = 1, face= "bold"), 
        axis.title.y = element_text(size = 14, face = "bold"), 
        legend.title = element_text(size = 16, face = "bold"),
        axis.title.x = element_text(size = 12, face = "bold"),
        legend.text = element_text(size = 14, face = "bold", colour = "black"),
        legend.position = "bottom",
        axis.text.y = element_text(colour = "black", size = 14, face = "bold"),
        plot.title = element_text(size=14,face="bold"),
        plot.caption = element_text(hjust = 0,vjust=0.5,size = 12, face= "bold", colour = "red"),
        panel.background = element_blank(),
        panel.border =element_rect(fill = NA, color = "black"), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"))
  


htmltools::HTML("<br>")
htmltools::HTML("<br>")


reads1[,8]<-data.frame(reads1$Cleaned/reads1$Raw)*100
colnames(reads1)[8]<-"Clean_Reads_Percent"
reads1[,8]<-round(reads1[,8],digits = 2)



datatable(reads1, 
          rownames= FALSE,
          filter = 'top',
          extensions = 'Buttons',
          options = list(scrollX = TRUE,
                         pageLength = 5,
                         dom = 'Blfrtip',
                         buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                         rownames = F),
                         caption = htmltools::tags$caption(style = 'caption-side: bottom; text-align: left;',
                         htmltools::em('Tracking Read Counts Across the DADA2 Pipeline:'), htmltools::br(), 
                         htmltools::em('Raw: Initial number of reads in each sample.'), htmltools::br(),
                         htmltools::em('Quality_Trimmed: Reads left after primer removal (if applicable) and quality trimming prior to DADA2 processing.'), htmltools::br(),
                         htmltools::em('Filtered: Reads left after eliminating sequences with ambiguous bases, high error rates, and short length.'), htmltools::br(),
                         htmltools::em('Denoised: Reads left with high quality and error-corrected sequences.'), htmltools::br(),
                         htmltools::em('Merged: Reads left after removal of singletons and unmerged forward and reverse read pairs.'), htmltools::br(),
                         htmltools::em('Cleaned: Reads left after eliminating chimeric sequences,artifact sequences created during PCR due to partially extended strands binding to similar but different sequences.')))


reads2<-data.frame(rep("Starting number of reads",times=nrow(reads1)),reads1$Raw)
colnames(reads2)<-c("Var","value")


reads3<-data.frame(rep("Finishing number of reads",times=nrow(reads1)),reads1$Cleaned)
colnames(reads3)<-c("Var","value")


```

<br>
<br>
<br>
<br>
<br>

## Distribution of initial and final read counts

<br>

```{r boxplot,echo=F,messages=FALSE, warning=FALSE,results = FALSE,fig.width=11,fig.height=7}

n=max(reads2$value)/10

# Determine max bin count
max_count <- max(hist(reads2$value, breaks=seq(0, max(reads2$value), by=n), plot=FALSE)$counts)

A<-ggplot(reads2, aes(x=value)) + 
  geom_histogram(color="black", fill="#E69F00",binwidth = n,boundary=0)+ #boundary start starts the bin from zero
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, size = 12, colour = "black",  hjust = 1, face= "bold"),
        axis.title.y = element_text(size = 12, face = "bold"), legend.title = element_text(size = 16, face = "bold"),
        axis.title.x = element_text(size = 12, face = "bold"),
        legend.text = element_text(size = 12, face = "bold", colour = "black"),
        axis.text.y = element_text(colour = "black", size = 12, face = "bold"),
        plot.title = element_text(size=14,face="bold"),
        plot.caption = element_text(hjust = 0,vjust=1.5,size = 12, face = "bold", colour = "red"),
        legend.position = "None")+xlab("Raw reads (count)")+ylab("Sample count")+
  ggtitle("Starting number of reads distribution")+
  theme(panel.background = element_blank(),
        panel.border =element_rect(fill = NA, color = "black"), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour = "black"))+
  # update: adds counts above bars
  stat_bin(binwidth = n,boundary = 0, aes(label=ifelse(..count.. > 0, ..count.., "")),geom="text",vjust=-0.9,hjust=0.5)+ 
coord_cartesian(ylim=c(0, max_count + 30))#+labs(caption="Binwidth is 5.")



n=max(reads3$value)/10

# Determine max bin count
max_count <- max(hist(reads3$value, breaks=seq(0, max(reads3$value), by=n), plot=FALSE)$counts)

B<-ggplot(reads3, aes(x=value)) +
  geom_histogram(color="black", fill="#E69F00",binwidth = n,boundary=0)+ #boundary start starts the bin from zero
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, size = 12, colour = "black",  hjust = 1, face= "bold"),
        axis.title.y = element_text(size = 12, face = "bold"), legend.title = element_text(size = 16, face = "bold"),
        axis.title.x = element_text(size = 12, face = "bold"),
        legend.text = element_text(size = 12, face = "bold", colour = "black"),
        axis.text.y = element_text(colour = "black", size = 12, face = "bold"),
        plot.title = element_text(size=14,face="bold"),
        plot.caption = element_text(hjust = 0,vjust=1.5,size = 12, face = "bold", colour = "red"),
        legend.position = "None")+xlab("Clean reads (count)")+ylab("Sample count")+
  ggtitle("Final number of clean reads distribution")+
  theme(panel.background = element_blank(),
        panel.border =element_rect(fill = NA, color = "black"), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        axis.line = element_line(colour ="black"))+
  # update: adds counts above bars
  stat_bin(binwidth = n,boundary = 0, aes(label=ifelse(..count.. > 0, ..count.., "")),geom="text",vjust=-0.9,hjust=0.5)+ 
coord_cartesian(ylim=c(0, max_count + 30))#+labs(caption="Binwidth is 5.")


ggarrange(A,B)

```

<br>
<br>
<br>
<br>
<br>

# **Investigating ASVs**

<br>

## Bacterial composition in the positive controls

<br>


```{r, warning=F, message=F, results='asis',echo=F,fig.width=10, fig.height=7}

#Read sequence data and taxonomy data
seqtab <- data.frame(read.csv(file = snakemake@params[["seqtab"]],row.names = 1))
taxa <- read.table(snakemake@params[["taxonomy"]], header = TRUE, sep = "\t", check.names = FALSE)

# Prepare the sample data
data <- data.frame(samples = rownames(seqtab))
rownames(data) <- data$samples


# Create a phyloseq object
ps <- phyloseq(otu_table(as.matrix(seqtab), taxa_are_rows = FALSE),
               sample_data(data),
               tax_table(as.matrix(taxa)))



# Prepare color palette
n <- 74
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))


# List of positive control samples
samples <- unlist(base::strsplit(snakemake@params[["pos"]], split = "\\|"))


if (length(samples) > 0 && !all(samples == "")) {
  
  # Define known bacterial community counts for each reference file
  ref_files <- list.files(snakemake@params[["ref"]])
  names(ref_files)<-stringr::str_split_fixed(ref_files, "\\.", 2)[, 1]
  
  #read in community composition of reference files in alphabetic order
  source(snakemake@params[["source"]])
  
  
  for (i in samples) {
    mock <- prune_samples(ps@sam_data$samples == i, ps)
    mock <- prune_taxa(colSums(mock@otu_table) > 0, mock)
    
    # Iterate over each reference file
    for (j in names(ref_files)) {
      mock.ref <- getSequences(paste0(snakemake@params[["ref"]],ref_files[[j]]))
      unqs.mock <- mock@otu_table@.Data
      
      # Retrieve the expected bacterial community counts for the current reference file
      expected_counts <- community_counts[[j]]
      
      match.ref <- as.data.frame(sapply(colnames(unqs.mock), function(x) grepl(x, mock.ref)))
      
      match.ref$species <- stringr::str_split_fixed(names(mock.ref), "_16", 2)[, 1]
      
      maps <- gather(data = match.ref, key = "seqs", value = "present", -species) %>%
        dplyr::filter(present == TRUE)

        #sometimes there is overlap between positive control reference files bacterial composition so samples get matched with a reference when there is >60% similar bacterial composition

        if(nrow(maps)>0 && mean(unique(match.ref$species) %in% unique(maps$species))*100 > 60){
        maps$abundance <- c(mock@otu_table[, maps$seqs])
        
        tot_abund<-maps %>% summarise(abundance=sum(abundance))
        
        temp=maps %>% dplyr::group_by(species) %>% dplyr::summarise("Observed"=round(100*(sum(abundance)/unlist(tot_abund)),digits = 1))
        colnames(temp)[1]<-"Taxa"
        
        temp<-left_join(expected_counts,temp,by="Taxa") %>%
        mutate(Observed = replace_na(Observed, 0))
               
        # Plotting
        A <- ggplot(reshape2::melt(as.data.frame(temp)), aes(x = variable, y = value, fill = Taxa)) +
          geom_bar(colour = "black", position = "stack", stat = "identity", width = 0.8) +
          xlab(NULL) +
          ggtitle(paste0("Bacterial profile in ", i)) +
          theme(legend.position = "right",
                axis.text.x = element_text(angle = 90, size = 15, colour = "black", vjust = 0.9, hjust = 0.83, face = "bold"),
                axis.title.y = element_text(size = 15, face = "bold"),
                axis.title.x = element_text(size = 15, face = "bold"),
                legend.title = element_text(size = 12, face = "bold"),
                legend.text = element_text(size = 12, face = "bold", colour = "black"),
                axis.text.y = element_text(colour = "black", size = 13, face = "bold"),
                plot.caption = element_text(hjust = 0, vjust = 0, size = 17, face = "bold", colour = "red"),
                plot.title = element_text(size = 15, face = "bold"),
                strip.text.x = element_text(size = 15, colour = "black")) +
          ylab("Counts") +
          scale_fill_manual(values = col_vector) +
          theme(panel.background = element_blank(),
                panel.border = element_rect(fill = NA, color = "black"),
                panel.grid.major = element_blank(),
                panel.grid.minor = element_blank(),
                axis.line = element_line(colour = "black"))+guides(fill=guide_legend(nrow=21,byrow=F))
        print(A)
        
        cat("<br><br>")
        
        B <- kable(temp, format = "html", caption = "Abundance table of taxa found in the positive control sample") %>%
          kable_styling() %>%
          column_spec(1:ncol(temp), width = "6cm")
        print(B)
        
        cat("<br><br>")
        cat("<br><br>")
        cat("<br><br>")
        cat("<br><br>")
        
      }
    }
  }
  
} else {
  cat("There was no positive control in this run")
  cat("<br><br>")
  cat("<br><br>")
  cat("<br><br>")
  cat("<br><br>")
  
}


```



## ASVs prevalence and Abudnance

<br>


```{r, investigating ASVs,echo=F,messages=FALSE, warning=FALSE,include=T}


seqtab <- data.frame(read.csv(file = snakemake@params[["seqtab"]],row.names = 1))

# Calculate prevalence for each taxa
Prevalence <- seqtab %>%
  summarise_all(~sum(. > 0)) %>%    # Prevalence: Count of non-zero entries for each taxa
  t() %>%
  as.data.frame() %>%
  mutate(Taxa = rownames(.))

colnames(Prevalence) <- c("Prevalence", "Taxa")

# Calculate abundance for each taxa
Abundance <- seqtab %>% 
  t() %>%
  rowSums() %>%  
  as.data.frame() 

Abundance$Taxa <- rownames(Abundance)

colnames(Abundance) <- c("Abundance", "Taxa")

# Joining prevalence and abundance
df <- left_join(Prevalence, Abundance, by = "Taxa")

# Melting the data for boxplot visualization
df1 <- reshape2::melt(df, id.vars = "Taxa", value.name = "count")

# Scatter plot
ggplot(df, aes(x = Abundance, y = Prevalence)) +
  geom_point(aes(color = ifelse(Prevalence > 10 & Abundance > 150, "#77c593", "black"))) +
  scale_color_identity() +
  ggtitle("Overview of ASVs focusing on their prevalence and abundance")+
  labs(caption = paste0("Points in green represent ASVs with prevalence > 10 and abundance > 150.\n",
  "Total number of ASVs is ", length(df$Taxa), " in ", nrow(seqtab) ," samples."))+
  theme(plot.caption = element_text(hjust = 0,size=12),
        axis.text.x = element_text(angle = 45, size = 12, colour = "black",  hjust = 1, face= "bold"),
        axis.title.y = element_text(size = 12, face = "bold"),
        legend.title = element_text(size = 16, face = "bold"),
        axis.title.x = element_text(size = 12, face = "bold"),
        legend.text = element_text(size = 12, face = "bold", colour = "black"),
        axis.text.y = element_text(colour = "black", size = 12, face = "bold"),
        plot.title = element_text(size=14,face="bold"),
        panel.background = element_blank(),
        panel.border =element_rect(fill = NA, color = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"))

```


<br>
<br>
<br>
<br>
<br>

## ASVs length distribution for all samples

<br>

```{r all ASVs length distribution abundance, results='asis', echo=F}

files <- list.files(path = snakemake@params[["length_distribution"]], pattern = "distribution", full.names =TRUE)


for (f in files) {
   cat(paste0("![](", f, "){width=1000%}\n"))
 }

```

<br>
<br>
<br>
<br>
<br>



## Taxonomy profile

<br>

```{r krona-plot, echo=FALSE,message=F, warning=F}

options(browser = "echo") #plot_krona is set to open the krona plot as a separate html file, this stops it

# Plot Krona chart
a <- plot_krona(physeq=ps, output=snakemake@params[['krona']], variable="samples", trim = TRUE)

file=snakemake@config[['path']]

```

<iframe src="`r paste0(file, "/output/QC_html_report/krona_Species_result.html")`" width="1100" height="700" scrolling="yes"></iframe>


<div Class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>
