---
title: "QC Run Quality Report"
author:
    - "IMC Bioinformatics"
date: "`r format(Sys.time(), '%d %B, %Y')`"
params:
   rmd: "qc_report.Rmd"
output:
  html_document:
  highlight: tango
  number_sections: no
  theme: default
  toc: yes
  toc_depth: 3
  toc_float:
    collapsed: no
    smooth_scroll: yes
---



## Overall read quality plots {.tabset .tabset-fade}

<br>

### Read quality raw 

<br>



```{r,results='asis', echo=F}
Sys.setenv(RSTUDIO_PANDOC="/home/sahar.bagheri1/softwares/miniconda/envs/rmd/bin")

files <- list.files(path = snakemake@params[["quality"]], pattern = "raw", full.names = TRUE)

for (f in files) {
  cat(paste0("![](", f, "){width=50%}"))
 }

```


### Read Quality after read trimming

<br>

```{r, results='asis',echo=F}

files <- list.files(path = snakemake@params[["quality"]], pattern = "afterQC", full.names =TRUE)

for (f in files) {
  cat(paste0("![](", f, "){width=50%}"))
 }
```


### Read Quality after read filtering

<br>

```{r, results='asis', echo=F}

files <- list.files(path = snakemake@params[["quality"]], pattern = "afterdada2", full.names =TRUE)

for (f in files) {
  cat(paste0("![](", f, "){width=50%}"))
 }
```


## Read length distribution for 5 random samples in the run {.tabset .tabset-fade}

<br>

### Sample 1

```{r, results='asis', echo=F}

files <- list.files(path = snakemake@params[["length_distribution"]], pattern = "S1", full.names =TRUE)

for (f in files) {
  cat(paste0("![](", f, ")\n"))
 }

```



### Sample 2

```{r, results='asis', echo=F}

files <- list.files(path = snakemake@params[["length_distribution"]], pattern = "S2", full.names =TRUE)

for (f in files) {
  cat(paste0("![](", f, ")\n"))
 }
```



### Sample 3

```{r, results='asis', echo=F}

files <- list.files(path = snakemake@params[["length_distribution"]], pattern = "S3", full.names =TRUE)

for (f in files) {
  cat(paste0("![](", f, ")\n"))
 }
```



### Sample 4

```{r, results='asis', echo=F}

files <- list.files(path = snakemake@params[["length_distribution"]], pattern = "S4", full.names =TRUE)

for (f in files) {
  cat(paste0("![](", f, ")\n"))
 }
```



### Sample 5

```{r, results='asis', echo=F}

files <- list.files(path = snakemake@params[["length_distribution"]], pattern = "S5", full.names =TRUE)

for (f in files) {
  cat(paste0("![](", f, ")\n"))
 }
```



<br>
<br>


```{r,echo=F,messages=FALSE, warning=FALSE}

suppressMessages(library(DT))
suppressMessages(library(dplyr))
suppressMessages(library(ggplot2))


reads<-read.csv(snakemake@params[["Nread"]],sep="\t") %>% filter(!grepl('Undetermined|undetermined', file))

#datatable(reads,caption="Read processing")

```

<br>
<br>

## Reads count throughout the dada2 pipeline in all samples

<br>

```{r table merge,echo=F,messages=FALSE, warning=FALSE,include=T,fig.width=7,fig.height=5}
reads1<-reads

#reads1[,8]<-data.frame(reads1$merged/reads1$num_seqs)*100
#colnames(reads1)[8]<-"Overlapped_Percent"
#reads1[,8]<-round(reads1[,8],digits = 2)


reads1[,8]<-data.frame(reads1$nonchim/reads1$num_seqs)*100
colnames(reads1)[8]<-"Clean_Reads_Percent"
reads1[,8]<-round(reads1[,8],digits = 2)


#reads2<-reads1 %>% filter(Clean_Reads_Percent>3)  #Check for the percentage threshold for excluding negative controls and undetermined rows, so far 40% has worked


datatable(reads1, 
          rownames= FALSE,
          filter = 'top',
          extensions = 'Buttons',
          options = list(scrollX = TRUE,
                         pageLength = 5,
                         dom = 'Blfrtip',
                         buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                         caption = "Read processing",
                         rownames = F)) 



reads2<-data.frame(rep("Starting number of reads",times=nrow(reads1)),reads1$num_seqs)
colnames(reads2)<-c("Var","value")


reads3<-data.frame(rep("Current run",times=nrow(reads1)),reads1$Clean_Reads_Percent)
colnames(reads3)<-c("Var","value")



```


```{r boxplot,echo=F,messages=FALSE, warning=FALSE,results = FALSE,include=F,fig.width=7,fig.height=5,eval=F}

ggplot(reads3,aes(x=Var,y = value,fill=Var))+geom_boxplot()+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, size = 12, colour = "black",  hjust = 1, face= "bold"),
        axis.title.y = element_text(size = 12, face = "bold"), legend.title = element_text(size = 16, face = "bold"),
        axis.title.x = element_text(size = 12, face = "bold"),
        legend.text = element_text(size = 12, face = "bold", colour = "black"),
        axis.text.y = element_text(colour = "black", size = 12, face = "bold"),
        plot.title = element_text(size=14,face="bold"),
        plot.caption = element_text(hjust = 0.3,vjust=1.5,size = 12, face = "bold", colour = "red"),
        legend.position = "None")+ylim(c(0,100))+xlab(NULL)+ylab("Clean reads")+
  ggtitle("Remained Clean reads percentage after dada2 in the current run")+theme(panel.background = element_blank(),panel.border =element_rect(fill = NA, color = "black"), panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = 
element_line(colour = "black"))


```

<br>
<br>
<br>
<br>
<br>

## Distribution of initial read counts

<br>

```{r raw reads distribution plot,echo=F,messages=FALSE, warning=FALSE,results = FALSE,include=T,fig.width=9,fig.height=5,eval=T}

n=max(reads2$value)/10

ggplot(reads2, aes(x=value)) + 
  geom_histogram(color="black", fill="#E69F00",binwidth = n,boundary=0)+ #boundary start starts the bin from zero
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, size = 12, colour = "black",  hjust = 1, face= "bold"),
        axis.title.y = element_text(size = 12, face = "bold"), legend.title = element_text(size = 16, face = "bold"),
        axis.title.x = element_text(size = 12, face = "bold"),
        legend.text = element_text(size = 12, face = "bold", colour = "black"),
        axis.text.y = element_text(colour = "black", size = 12, face = "bold"),
        plot.title = element_text(size=14,face="bold"),
        plot.caption = element_text(hjust = 0,vjust=1.5,size = 12, face = "bold", colour = "red"),
        legend.position = "None")+xlab("Raw reads (count)")+ylab("Sample count")+
  ggtitle("Starting number of reads distribution")+theme(panel.background = element_blank(),panel.border =element_rect(fill = NA, color = "black"), panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = 
"black"))+
  # update: adds counts above bars
  stat_bin(binwidth = n,boundary = 0, aes(label=ifelse(..count.. > 0, ..count.., "")),geom="text",vjust=-0.9,hjust=0.5)#+labs(caption="Binwidth is 5.")

#scale_x_continuous(breaks = seq(0,400000,))+ylim(c(0,60))

```

<br>
<br>
<br>
<br>
<br>

## Percentage of clean reads post-DADA2

<br>

```{r distribution plot,echo=F,messages=FALSE, warning=FALSE,results = FALSE,include=T,fig.width=9,fig.height=5,eval=T}

ggplot(reads3, aes(x=value)) + 
  geom_histogram(color="black", fill="#E69F00",binwidth = 5,boundary=0)+   #boundary start starts the bin from zero
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, size = 12, colour = "black",  hjust = 1, face= "bold"),
        axis.title.y = element_text(size = 12, face = "bold"), legend.title = element_text(size = 16, face = "bold"),
        axis.title.x = element_text(size = 12, face = "bold"),
        legend.text = element_text(size = 12, face = "bold", colour = "black"),
        axis.text.y = element_text(colour = "black", size = 12, face = "bold"),
        plot.title = element_text(size=14,face="bold"),
        plot.caption = element_text(hjust = 0,vjust=1.5,size = 12, face = "bold", colour = "red"),
        legend.position = "None")+xlab("Clean reads (%)")+ylab("Sample count")+
  ggtitle("Remained Clean reads percentage after dada2 in the current run")+theme(panel.background = element_blank(),panel.border =element_rect(fill = NA, color = "black"), panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = 
element_line(colour = "black"))+
  # update: adds counts above bars
  stat_bin(binwidth = 5,boundary = 0, aes(label=ifelse(..count.. > 0, ..count.., "")),geom="text",vjust=-0.9)#+labs(caption="Binwidth is 5.")

#  scale_x_continuous(breaks = seq(0,100,10))+ylim(c(0,70))

```

## Source
<a download="qc_report.Rmd" href="`r base64enc::dataURI(file = params$rmd, mime = 'text/rmd', encoding = 'base64')`">R Markdown source file (to produce this document)</a>

